---
title: "R Notebook"
output: html_notebook
---

This is a [R Notebook](http://rmarkdown.rstudio.com) for the "Introducton to R and R Studio" UMN Data Management Bootcamp workshop. Slides and data can be found [here](http://z.umn.edu/rtutorial). 

# Objectives
		- Exploring RStudio
		- Setting up a new project
		- Getting data into R as a CSV file
		- Exploring your data & getting summary statistics
		- Plotting Your Data
		- Get data out of R as a CSV file

***

# Aside: What is an RNotebook (i.e. "What Am I Looking At?)

RNotebooks let you run R code and display the results while also sharing the results as an html file. We will be using this R Notebook to display code and examples. All the commands seen in the following snippits can be run in an R Script, the R console, or in a .R file

For example the following displays a simple scatterplot for the data in the built-in dataset "cars" (which contains data about cars!) for the variable "mpg". This runs just as it would in a R Script but it also generates a .html file with the output. The file for this R Notebook is located [here](https://fdsayre.github.io/TeachingRIntro50Minutes/50MinuteRTutorial.nb.html)

```{r}
plot(mtcars$mpg)
```


# Exploring RStudio

Things to point out:
- File editing pane
- Console/Terminal pane
- Right panes (enviornment, information)
- How to change settings (and themes)

More information we are going to skip over:
- [Data Carpentry: Before We Start](http://datacarpentry.org/R-ecology-lesson/00-before-we-start.html) 



***

# Seting up a new project

Using "File --> New Project" set up a new project in a new directory on your computer. This is where we will save our work (remember where this is so we can move data to it later). 

## Create a new R Script

Open a new R Script file (File --> New File --> R Script) and save it in your working directory with a .R extention

## Set your working directory

This lets R know where it should look for and save files

This can be done by (Session --> Set Working Directory). 

Then copy and paste the line that appears in your console to your script so it's easy to run next time

```{r}
setwd("~/projects/TeachingRIntro50Minutes")
```

## Install any libraries you will need

This only needs to be done once for each install of R. This can take a lot of time if the package is large or has a lot of dependencies. In this case, let's start by installing the ggplot2 package. 

```{r}
install.packages("ggplot2")
```


## Load Libraries into memory. 

This needs to be done each for each R session. 

```{r}
library(ggplot2)
```


## ASIDE: Getting help inside R for any of the above commands

```{r}
?str
?read.csv
```


***

# Getting data into R

There a number of different ways to get data into R. 

## Manually assigning variables

First, let's assign some data, both numbers and text, to a couple different variable names. 

ASIDE: R has a lot of different data structures and types that you will learn as you explore R and as you need to do different things with R.

```{r}
x <- 5
x
y <- "Rstats"
y
z <- c("rstats", "perl", "bash", "python")
z
```


## Built-in dataset that comes with R 
R comes with a number or built-in datasets that are good for eploring R and R Studio. We can load them directly but we can also assign them to variable names.


```{r}
mtcars
cardata <- mtcars
cardata
```

# Loading our own dataset (from .csv files) into R
the read.csv function will load a csv file into RStudio. It has a number of options. 

```{r}
data <- read.csv("mpls_energy_benchmarking_2015.csv", header=TRUE, na.strings=c("N/A", "Not Available", "NA", "0"))

```

You can also download data directly from R, in this case from figshare: 

```{r}
download.file("https://ndownloader.figshare.com/files/2292169", "portal_data_joined.csv")

surveys <- read.csv("portal_data_joined.csv")

```

***

# Exploring your data

## Look at the first couple rows of your data

```{r}
head(data)
```

## See the structure of your data

```{r}
str(data)
```


## See the number of rows and columns in your data

```{r}
nrow(data)
ncol(data)
```

## Basic summary statistics

```{r}
summary(data$energy_star_score)
```


***

# Plotting your data

We can use the built-in plotting library that comes with base R to plot our data in various ways: 

```{r}
plot(data$energy_star_score)
```


```{r}
hist(data$energy_star_score)
```

```{r}
boxplot(data$energy_star_score)
```

```{r}
boxplot(data$energy_star_score ~ data$org_name)
```

Where missing some labels! Lets try this again! (where did I find this information? I googled it: https://www.statmethods.net/advgraphs/parameters.html)

```{r}
boxplot(data$energy_star_score ~ data$org_name, las=2, cex.axis=.5)
```

## GGPlot

Let's look at the Tidyverse GGPLOT package using the mtcars built-in dataset we assigned to the variable "cardata" earlier

- [GGPlot](https://ggplot2.tidyverse.org/reference/ggplot.html)
- [GGPlot Cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf)

```{r}

ggplot(cardata, aes(y=mpg,x=hp)) +
  geom_point()

```

```{r}

ggplot(cardata, aes(y=mpg,x=cyl)) +
  geom_bar(stat="identity")

```

What else should we do?

***

# Getting data out of R

## Export a dataset to a CSV file

```{r}
write.csv(data, "data.csv")
```

